---
title: "Redownload data"
author: "sbsambado"
date: "2022-12-15"
output: html_document
---

# Manuscript - "TITLE"
## Author(s) - S Sambado, A J MacDonald, A Swei, C J Briggs

### Script 1. Data cleaning

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## necessary packages
library(readr)
library(tidyverse)
library(ggplot2)
library(lubridate)

## raw datasets needed 

# from personal collections
tick_full <- read.csv("data/California Tick Data Base - Sheet1_20221215.csv")
liz_tidy <- read.csv("Mathias2021_FieldData_Lizard - lizard_tidy_20221215.csv")
liz_wide <- read.csv("Mathias2021_FieldData_Lizard - lizard_long_20221215.csv")

# from remote sensed data
prism <- read.csv("data/PRISM_FULL_reserves.csv")


```

# Datasets

*primer datasets*
+ `collection_dates.csv` : dataframe of collection days 

+ `prism_full_clean.csv` : dataframe of all climate variables (including sum ppt & ppt prior month) for collection days


**Main dataframes for analyses**
+ `drag_climate_lifestage` : dataframe of aggregated tick counts by reserve, month and life stage from drag cloths with climate data on collection day

+ `liz_climate_lifestage` : dataframe of aggregated tick counts on individual lizards by reserve, month and each life stage (so two entries for one lizard per month and reserve) with climate data on collection day

+ `liz_wide_climate_prop` : dataframe of nymphal proportion index and binary outcomes (1 = synchronous feeding, 0 = not synchronous feeding) on individual lizards with climate data

*Other dataframes created*
+ `drag_climate_nolifestage` : dataframe of aggregated tick counts by reserve and month from drag cloths with climate data on collection day

+ `liz_climate_lifestage` : dataframe of aggregated tick counts on individual lizards by reserve and month with climate data on collection day




# Format climate data

# make collection day

```{r}

tick_full <- read.csv("data/California Tick Data Base - Sheet1_20221215.csv")

tick_sub <- tick_full %>%
    group_by(reserve, month) %>%
  filter(project == "phenology" & genus == "Ixodes" ) 

tick_clean <- subset(tick_sub, select = c(1,4:5, 10,17,18))

collection_date <- tick_clean %>%
  group_by(reserve, month) %>%
  dplyr::select(collection_date) %>%
  slice(1)

# make data compatible 
collection_date$collection_date <- mdy(collection_date$collection_date)

# updated bug i found: change Coal Oil Point collection_date from 2021-03-15 to 2021-05-15
collection_date$collection_date[collection_date$reserve == "Coal Oil Point" & collection_date$collection_date == "2021-03-15"] <- "2021-05-15"

# check to make sure it worked
collection_date %>%
  filter(reserve == "Coal Oil Point")

names(collection_date)[2] <- "month_name"
collection_date$month <- substr(collection_date$collection_date,6,7)
collection_date$month <- as.numeric(collection_date$month)



write.csv(collection_date, "data/collection_date.csv")

```

# match prism data to collection data

step 1. spread prism & formulate dates
`prism_spread` is final product
```{r}
# pull in full prism data

prism <- read.csv("data/PRISM_FULL_reserves.csv")

prism_sub <- subset(prism, select = c(3:6))

# spread data

prism_spread <- prism_sub %>%
  spread(key = data_type, value = data)

# make date lubridated
prism_spread$date <- ymd(prism_spread$date)

# make a month column

prism_spread$month <- substr(prism_spread$date, 6,7)
prism_spread$month <- as.numeric(prism_spread$month) # make sure to match value from other datasets
```


step 2. create ppt values for ppt per month and ppt prior month

`prism_ppt` is final product
```{r}
# hold on this is a dumb and long way to go about this


# make column for sum of ppt per month
prism_spread_pptmonth <- prism_spread %>%
  group_by(reserve, month) %>%
  mutate(ppt_month = sum(ppt))

# make column for sum of PREVIOUS month
reserve_ppt_month <- prism_spread_pptmonth %>%
  group_by(reserve, month) %>%
  dplyr::select(ppt_month) %>%
  slice(1)

# give month numeric value
reserve_ppt_month$month <- as.numeric(reserve_ppt_month$month)

# make fake month 
reserve_ppt_month$month_plus1 <- reserve_ppt_month$month + 1

# now change 13 to 1 to make it ordered
reserve_ppt_month$month_plus1[reserve_ppt_month$month_plus1 == 13] <- 1 

# reorder fake months & then calculate lag

prism_ppt <- reserve_ppt_month %>%
  group_by(reserve) %>%
  arrange(month_plus1) %>%
dplyr::mutate(ppt_prior_month = lag(ppt_month, n = 1, default = NA)) %>%
  #filter(month != 12 & month  != 7) %>%
  dplyr:: select("reserve", "month", "ppt_month", "ppt_prior_month")



```

now add ppt data to prism spread
```{r}
prism_spread
prism_ppt

full_prism <- merge(prism_spread, prism_ppt, by = c("reserve", "month"))
```


now lets only select for values on collection dates
```{r}
full_prism # 2,343 × 11
collection_date # 47 × 4

# format to tibbles for easier manipulation
full_prism <- as.tibble(full_prism)

# change name to match for merging
names(collection_date)[3] <- "date" 

# want to have just collection date data

prism_full <- left_join(collection_date, full_prism, by = c("reserve","date"))

# lets clean this up for just the stuff we want and labels

prism_clean <- subset(prism_full, select = c(1,2,3,4,6:13))

names(prism_clean)[4] <- "month_numeric"


## add all the reserve extras

prism_clean$region <- 1

prism_clean$region[prism_clean$reserve == "McLaughlin"] <- "northern"
prism_clean$region[prism_clean$reserve == "Quail Ridge"] <- "northern"
prism_clean$region[prism_clean$reserve == "Point Reyes"] <- "northern"

prism_clean$region[prism_clean$reserve == "Fort Ord"] <- "central"
prism_clean$region[prism_clean$reserve == "Hastings"] <- "central"
prism_clean$region[prism_clean$reserve == "Big Creek"] <- "central"
prism_clean$region[prism_clean$reserve == "Rancho Marino"] <- "central"

prism_clean$region[prism_clean$reserve == "Sedgwick"] <- "southern"
prism_clean$region[prism_clean$reserve == "Coal Oil Point"] <- "southern"
prism_clean$region[prism_clean$reserve == "Stunt Ranch"] <- "southern"
prism_clean$region[prism_clean$reserve == "Santa Cruz Island"] <- "southern"


prism_clean$reserve <- factor(prism_clean$reserve,
                      levels = c("McLaughlin", "Quail Ridge", "Point Reyes",
                                 "Fort Ord", "Hastings", 
                                 "Big Creek", "Rancho Marino",
                                 "Sedgwick", "Coal Oil Point",
                                 "Stunt Ranch", "Santa Cruz Island"))



prism_clean$location <- prism_clean$region

prism_clean$location[prism_clean$reserve == "McLaughlin"] <- "inland"
prism_clean$location[prism_clean$reserve == "Quail Ridge"] <- "inland"
prism_clean$location[prism_clean$reserve == "Point Reyes"] <- "coastal"

prism_clean$location[prism_clean$reserve == "Fort Ord"] <- "coastal"
prism_clean$location[prism_clean$reserve == "Hastings"] <- "inland"
prism_clean$location[prism_clean$reserve == "Big Creek"] <- "coastal"
prism_clean$location[prism_clean$reserve == "Rancho Marino"] <- "coastal"

prism_clean$location[prism_clean$reserve == "Sedgwick"] <- "inland"
prism_clean$location[prism_clean$reserve == "Coal Oil Point"] <- "coastal"
prism_clean$location[prism_clean$reserve == "Stunt Ranch"] <- "inland"
prism_clean$location[prism_clean$reserve == "Santa Cruz Island"] <- "inland"

prism_clean$region <- factor(prism_clean$region, 
                     levels = c("northern", "central", "southern"))

prism_clean$location <- as.factor(prism_clean$location)


unique(prism_clean$month_name)

prism_clean$month_name[prism_clean$month_name == "May "] <-  "May" 


prism_clean$month_name <- factor(prism_clean$month_name,
                                 levels = c("January", "March", "April", "May", "June"))
write.csv(prism_clean, "data/prism_full_clean.csv")


```

# format tick data

`tick_aggregate_lifestage` is product that lists total ticks per reserve per month by each life stage
`tick_aggregate_nolifestage` is product that lists just total ticks per reserve per month
```{r}
tick_full <- read.csv("data/California Tick Data Base - Sheet1_20221215.csv")

tick_sub <- tick_full %>%
    group_by(reserve, month) %>%
  filter(project == "phenology" & genus == "Ixodes" ) 

tick_clean <- subset(tick_sub, select = c(1,4:5, 10,17,18))

# add count 
tick_clean$count <- 1

# format month columns
names(tick_clean)[2] <- "month_name"
names(tick_clean)[3] <- "date"

# format date
tick_clean$date <- mdy(tick_clean$date )
# make sure all life stages are L, N, or A
unique(tick_clean$life_stage)

# convert ID life stage to N
tick_clean$life_stage[tick_clean$life_stage == "ID"] <- "N"
unique(tick_clean$life_stage)

tick_clean

tick_aggregate_lifestage <- aggregate(count ~ reserve + date + life_stage, FUN = sum, data = tick_clean) # 81, 3

tick_aggregate_nolifestage <- aggregate(count ~ reserve + date, FUN = sum, data = tick_clean)

```

# merge tick data with climate data
```{r}

## no life stage data
tick_aggregate_nolifestage
prism_clean

# need to make column to match prism_clean
tick_aggregate_nolifestage$month_numeric <- substr(tick_aggregate_nolifestage$date, 6,7)
tick_aggregate_nolifestage$month_numeric <- as.numeric(tick_aggregate_nolifestage$month_numeric)


drag_climate_nolifestage  <-left_join(tick_aggregate_nolifestage, prism_clean, by = c("reserve", "month_numeric")) %>%
  dplyr::select(-date.y) 

names(drag_climate_nolifestage)[2] <- "date" 

 
write.csv(drag_climate_nolifestage, "data/drag_climate_nolifestage.csv")


## with life stage
tick_aggregate_lifestage
prism_clean

# need to make column to match prism_clean
tick_aggregate_lifestage$month_numeric <- substr(tick_aggregate_lifestage$date, 6,7)
tick_aggregate_lifestage$month_numeric <- as.numeric(tick_aggregate_lifestage$month_numeric)

drag_climate_lifestage  <-left_join(tick_aggregate_lifestage, prism_clean, by = c("reserve", "month_numeric")) %>%
  dplyr::select(-date.y) 

names(drag_climate_lifestage)[2] <- "date" 
write.csv(drag_climate_lifestage, "data/drag_climate_lifestage.csv")


```

# tick
reserve, month (n = 53 groupings)
# now lizard data 
reserve, month, individual lizards (n = 297?)

Lizard tidy for neg binomial analyses
```{r}
liz_tidy <- read.csv("Mathias2021_FieldData_Lizard - lizard_tidy_20221215.csv")

# format data
liz_tidy$month_numeric <- substr(liz_tidy$date, 6,7)
liz_tidy$month_numeric <- as.numeric(liz_tidy$month_numeric)

liz_tidy$date <- ymd(liz_tidy$date)

# merge with prism
# with life stage 
liz_climate_lifestage  <-left_join(liz_tidy, prism_clean, by = c("reserve", "month_numeric")) %>%
  dplyr::select(-c(date.y, round)) 
names(liz_climate_lifestage)[2] <- "date"
liz_climate_lifestage$lifestage <- as.factor(liz_climate_lifestage$lifestage)

liz_climate_lifestage <-liz_climate_lifestage[1:594,]

write.csv(liz_climate_lifestage, "data/liz_climate_lifestage.csv")

# no life stge 


liz_tidy_nolifestage <- liz_tidy %>%
  group_by(reserve, month_numeric, lizard_num) %>%
  filter(reserve != "") %>%
  summarise(count = sum(count))


unique(liz_tidy_nolifestage$reserve) #"Coal Oil Point"
unique(prism$reserve) #               "Coal Oil Point"

liz_tidy_nolifestage %>%
  filter(reserve == "Coal Oil Point") 

prism_clean %>%
  filter(reserve == "Coal Oil Point")  # so COPR has month_numeric for May because date was put in as 2021-03-15 instead of 2021-05-15

# change that 

liz_climate_nolifestage  <-left_join(liz_tidy_nolifestage, prism_clean, by = c("reserve", "month_numeric"))  # not sure why COPR is not getting merged


write.csv(liz_climate_nolifestage, "data/liz_climate_nolifestage.csv")

```

Lizard wide for proportion analyses

```{r}

liz_wide <- read.csv("Mathias2021_FieldData_Lizard - lizard_long_20221215.csv")

liz_wide_select <- liz_wide[,c(1,3,5,14)]

# format data

liz_wide_select$date <- mdy(liz_wide_select$date)

liz_wide_select$month_numeric <- substr(liz_wide_select$date, 6,7)
liz_wide_select$month_numeric <- as.numeric(liz_wide_select$month_numeric)
# add binned proportions

liz_wide_select_prop <- liz_wide_select %>%
  mutate(binned_prop = case_when(nymphal_propption <.3 ~ 0,
                                 nymphal_propption >= .3 & nymphal_propption <= .7 ~ 1,
                                 nymphal_propption > .7 ~ 0))

# merge with prism data
liz_wide_climate_prop  <-left_join(liz_wide_select_prop, prism_clean, by = c("reserve", "month_numeric"))  %>%
  dplyr::select(-date.y)

names(liz_wide_climate_prop)[2] <- "date" 


liz_wide_climate_prop

write.csv(liz_wide_climate_prop, "data/liz_wide_climate_prop.csv")


```

