---
title: "2_DataAnlyses_v3"
author: "Samantha Sambado"
date: "2024-05-29"
output: html_document
---

#Manuscript - Climate associated variations in the within-season dynamics of juvenile ticks in California

##Authors - S Sambado, A J MacDonald, A Swei, C J Briggs

###Script 2. Data analyses

#Main Text
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## upload necessary packages
library(readr)
library(ggplot2)
library(tidyverse)
library(rstatix)
library(MASS)
library(lme4)
library(psych)
library(aod)
library(car)
library(ggridges)
library(ggpubr)
library(ggbreak)

## upload necessary data
# data structures good for plotting
drag_climate_lifestage <- read.csv("data/drag_climate_lifestage.csv")
liz_climate_lifestage <- read.csv("data/liz_climate_lifestage.csv")
liz_wide_climate_prop <- read.csv("data/liz_wide_climate_prop.csv")

# data structures good for analysis
drag_climate_nolifestage <- read.csv("data/drag_climate_nolifestage.csv")
liz_climate_nolifestage <- read.csv("data/liz_climate_nolifestage.csv")


## make sure factors appropriate
drag_climate_lifestage$life_stage <- factor(drag_climate_lifestage$life_stage,
                                            levels = c("L", "N", "A"),
                                            labels = c("Larva", "Nymph", "Adult")) 


drag_climate_lifestage$month_name <- factor(drag_climate_lifestage$month_name, 
                                            levels = c("March", "April", "May", "June"))

liz_climate_lifestage$month_name <- factor(liz_climate_lifestage$month_name, 
                                            levels = c("March", "April", "May", "June"))



## additional edits to the data
# make density
drag_climate_lifestage$density_100 <- drag_climate_lifestage$count/15 # bc 1500 sampled per reserve per month


```


###Data Summary

```{r}
## dragged data

drag_total <- drag_climate_lifestage %>%
  tally(count) # 895

drag_climate_lifestage %>%
  group_by(life_stage) %>%
  tally(count)  # L = 253, N = 89, A = 553

## lizard data

liz_total <- liz_climate_lifestage %>%
  tally(count) # 1503	

liz_climate_lifestage %>%
  group_by(lifestage) %>%
  tally(count) # L = 713, N = 791


## ixodes total
drag_total + liz_total # 2399
```

###Question 1

(a) What are the regional within-season patterns of tick abundances? (b) Are monthly abundance patterns similar for question (i.e., drag cloth) and attached ticks (i.e., lizard burden) collections methods?


```{r}
###################### (a) Within-season patterns of tick abundances

#### Juvenile ticks on lizards

### Regional patterns - kruskal-wallis
# test
liz_climate_lifestage %>% 
  kruskal.test(count ~ region)

# post-hoc
liz_climate_lifestage %>%
  dunn_test(count ~ region) 

### Location patterns - wilcoxon
# test
wilcox.test(count ~ location, liz_climate_lifestage) 

# means
liz_climate_lifestage %>%
  group_by(location) %>%
  summarise(mean = mean(count), sd = sd(count))


### Effect sizes
## Entire study
# Regional patterns
liz_climate_lifestage %>% kruskal_effsize(count ~ region)
# Location patterns
liz_climate_lifestage %>% cohens_d(count ~ location, var.equal = FALSE)

## By month
## Regional patterns
#liz_climate_lifestage %>%  filter(month_name == "March") %>% kruskal_effsize(count ~ region)
liz_climate_lifestage %>%  filter(month_name == "April") %>% kruskal_effsize(count ~ region)
liz_climate_lifestage %>%  filter(month_name == "May") %>% kruskal_effsize(count ~ region)
liz_climate_lifestage %>%  filter(month_name == "June") %>% kruskal_effsize(count ~ region)

## Location patterns
liz_climate_lifestage %>%  filter(month_name == "March") %>% cohens_d(count ~ location, var.equal = FALSE)
liz_climate_lifestage %>%  filter(month_name == "April") %>% cohens_d(count ~ location, var.equal = FALSE)
liz_climate_lifestage %>%  filter(month_name == "May") %>% cohens_d(count ~ location, var.equal = FALSE)
liz_climate_lifestage %>%  filter(month_name == "June") %>% cohens_d(count ~ location, var.equal = FALSE)

```


```{r}
###################### (b) Monthly abundance patterns

#### Drag cloth ticks

### all life stages means
drag_climate_lifestage %>%
   group_by(month_numeric) %>%
   summarize(Mean = mean(count, na.rm = TRUE),
             Sd = sd(count, na.rm = TRUE))

### individual life stage means
drag_climate_lifestage %>%
   group_by(month_numeric, life_stage) %>%
  filter(life_stage == "L") %>%
   summarize(Mean = mean(count, na.rm = TRUE),
             Sd = sd(count, na.rm = TRUE)) 

drag_climate_lifestage %>%
   group_by(month_numeric, life_stage) %>%
  filter(life_stage == "N") %>%
   summarize(Mean = mean(count, na.rm = TRUE),
             Sd = sd(count, na.rm = TRUE)) 


drag_climate_lifestage %>%
   group_by(month_numeric, life_stage) %>%
  filter(life_stage == "A") %>%
   summarize(Mean = mean(count, na.rm = TRUE),
             Sd = sd(count, na.rm = TRUE)) 



#### Juvenile ticks on lizards

### all life stages means
liz_climate_lifestage %>%
   group_by(month_numeric) %>%
   summarize(Mean = mean(count, na.rm = TRUE),
             Sd = sd(count, na.rm = TRUE))

### individual life stage means
liz_climate_lifestage %>%
   group_by(month_numeric) %>%
  filter(lifestage == "larval") %>%
   summarize(Mean = mean(count, na.rm = TRUE),
             Sd = sd(count, na.rm = TRUE)) 

liz_climate_lifestage %>%
   group_by(month_numeric) %>%
  filter(lifestage == "nymphal") %>%
   summarize(Mean = mean(count, na.rm = TRUE),
             Sd = sd(count, na.rm = TRUE)) 


```


###Question 2
How does within-season tick abundances respond to temperature (north to south) and/or humidity (coastal to inland) gradients?

```{r}
###################### lizard - larvae
liz_climate_lifestage_larv <- liz_climate_lifestage %>%
  filter(lifestage == "larval")


## full model
fit_liz_3_1 <- glm.nb(count ~ tmean + tmin + vpdmax + ppt_month,
                      data = liz_climate_lifestage_larv)

summary(fit_liz_3_1) 

vif(fit_liz_3_1)

## slimmed model
fit_liz_3_2 <- glmer.nb(count ~ tmean + vpdmax  + (1|reserve), # less collinear of a model
                      data = liz_climate_lifestage_larv)

summary(fit_liz_3_2) 


vif(fit_liz_3_2)


###################### lizard - nymphs
liz_climate_lifestage_nymp <- liz_climate_lifestage %>%
  filter(lifestage != "larval")

## full model 
fit_liz_4_0 <- glm.nb(count ~ tmean + tmin + vpdmax + ppt_month,
                      data = liz_climate_lifestage_nymp)
summary(fit_liz_4_0)

fit_liz_4_1 <- glm.nb(count ~ tmean + vpdmin,  # tmax + 
                      data = liz_climate_lifestage_nymp)

summary(fit_liz_4_1)
vif(fit_liz_4_1)



## slimmed model
fit_liz_4_2 <- glmer.nb(count ~   tmean + vpdmin + (1|reserve),  # less collinear of a model
                      data = liz_climate_lifestage_nymp)

summary(fit_liz_4_2)

```




###Question 3
Do larvae and nymphs feed synchronously on individual hosts? If so, how frequently, and are there climatic variables associated with synchronous feeding patterns?


```{r}

## step 1. format data 
# make synchrony outcome
sync_data <- liz_climate_lifestage %>% 
  filter(count >=0) %>% 
  pivot_wider(names_from = lifestage, values_from = count) %>% 
  group_by(reserve, lizard_num, month_numeric) %>% 
  summarise(larvae = sum(larval, na.rm = T),
            nymphs = sum(nymphal, na.rm = T)) %>% 
  mutate(sync = 
           case_when(
             larvae >0 & nymphs > 0 ~ 1,
             TRUE ~ 0
           ))
sync_data$sync <- factor(sync_data$sync) 



## merge sync data with climate
# make just important climate with lizard data
liz_climate <- liz_climate_lifestage %>% 
  group_by(reserve,lizard_num, month_numeric) %>% 
  dplyr::select(ppt, tmax, tmean, tmin, vpdmax, vpdmin, ppt_month, ppt_prior_month) %>% 
  unique()

# merge both data sets
sync_climate <- merge(sync_data, liz_climate, by = c("reserve", "lizard_num", "month_numeric"))

# make sure data type is correct
sync_climate$reserve <- factor(sync_climate$reserve,
                               levels = c("Stunt Ranch",
                               "McLaughlin", "Quail Ridge", "Point Reyes",
                                  "Fort Ord", "Hastings", 
                                  "Big Creek", "Rancho Marino",
                                   "Santa Cruz Island",
                                  "Sedgwick", "Coal Oil Point"))

# remove COP because there were only nymphs attached to lizards
sync_climate2 <- sync_climate %>% 
  filter(reserve != "Coal Oil Point")


```


```{r}

## step 2. run models
logit <- glm(sync ~ tmean + reserve, data = sync_climate2, family = "binomial")
summary(logit)
# every 1 unit change in tmean, the log odds of synchrony decreases by -0.1277

# calculate CIs
confint(logit)
# CIs using profiled log-likelihood

# odds ratio and 95% CI
exp(cbind(OR = coef(logit), confint(logit)))

# with 1 unit incrase in tmean, odds of sync increase by a factor of 0.88
exp(.88)


### odds ratio = -0.12771
exp(-0.12771) # 0.8801086
exp(coef(logit))
exp(logit$coefficients)/(1+exp(logit$coefficients))



# overall effect of reserve using wald.test
wald.test(b = coef(logit), Sigma = vcov(logit), Terms = 3:10)
# X2 = 31.6, df = 8, P(> X2) = 0.00011
# The chi-square test statistics of 31.6 with 9 degrees of free is associated with p-value of 0.00011 indicating the the overall effect of reserve is statistically significant

```
#Supporting Information

##Covariate correlations
```{r}
# check correlation
climate_sub <- drag_climate_nolifestage %>%
  dplyr::select( ppt:ppt_prior_month)

pairs.panels(climate_sub)
```

##Fitting tick-climate models
```{r}
###################### Drag cloth method

## Drag -- Adult

drag_climate_lifestage_adult <- drag_climate_lifestage %>%
  filter(life_stage == "A")

## full model

fit_nb_adult_1 <- glm.nb(count ~ tmax + tmean + tmin + vpdmax + vpdmin +ppt_month + ppt_prior_month,
                   data = drag_climate_lifestage_adult)

summary(fit_nb_adult_1) # vpdmin > vpdmax

stepAIC(fit_nb_adult_1)

## slimmed model
fit_nb_adult_3 <- glmer.nb(count ~ tmean + vpdmax + (1|reserve), # build less collinear model
                         data = drag_climate_lifestage_adult)
summary(fit_nb_adult_3)


### Drag -- Nymph

drag_climate_lifestage_nymph <- drag_climate_lifestage %>%
  filter(life_stage == "N")

## full model

fit_nb_nymph_1 <- glm.nb(count ~ tmax + tmean + tmin + vpdmax + vpdmin +ppt_month + ppt_prior_month,
                   data = drag_climate_lifestage_nymph)

summary(fit_nb_nymph_1) # 

stepAIC(fit_nb_nymph_1)

## slimmed model
fit_nb_nymph_3 <- glmer.nb(count ~ ppt_month + (1|reserve), # build less collinear model
                         data = drag_climate_lifestage_nymph)
summary(fit_nb_nymph_3)


### Drag -- Larvae

drag_climate_lifestage_larva <- drag_climate_lifestage %>%
  filter(life_stage == "L")

## full model

fit_nb_larva_1 <- glm.nb(count ~ tmax + tmean + tmin + vpdmax + vpdmin +ppt_month + ppt_prior_month,
                   data = drag_climate_lifestage_larva)

summary(fit_nb_larva_1) # ppt prior month

step(fit_nb_larva_1)

## slimmed model
fit_nb_larva_2 <- glmer.nb(count ~ ppt_prior_month + (1|reserve), # build less collinear model
                         data = drag_climate_lifestage_larva)
summary(fit_nb_larva_2)

###################### Lizard burden method

## Lizard -- Larvae


liz_climate_lifestage_larv <- liz_climate_lifestage %>%
  filter(lifestage == "larval")


## full model
fit_liz_3_1 <- glm.nb(count ~ tmean + tmin + vpdmax + ppt_month,
                      data = liz_climate_lifestage_larv)

summary(fit_liz_3_1) 

vif(fit_liz_3_1)

stepAIC(fit_liz_3_1)

## slimmed model

fit_liz_3_2 <- glmer.nb(count ~ tmean + vpdmax  + (1|reserve), # less collinear of a model
                      data = liz_climate_lifestage_larv)

summary(fit_liz_3_2) 





## Lizard -- Nymph


liz_climate_lifestage_nymp <- liz_climate_lifestage %>%
  filter(lifestage != "larval")

## full model 
fit_liz_4_0 <- glm.nb(count ~ tmean + tmin + vpdmax + ppt_month,
                      data = liz_climate_lifestage_nymp)
summary(fit_liz_4_0)

stepAIC(fit_liz_4_0)

fit_liz_4_1 <- glm.nb(count ~ tmean + vpdmin,  # tmax + 
                      data = liz_climate_lifestage_nymp)

summary(fit_liz_4_1)
vif(fit_liz_4_1)

## slimmed model
fit_liz_4_2 <- glmer.nb(count ~   tmean + vpdmin + (1|reserve),  # less collinear of a model
                      data = liz_climate_lifestage_nymp)

summary(fit_liz_4_2)



```










