---
title: "3_Figures"
author: "sbsambado"
date: "2022-12-18"
output: html_document
---

# Manuscript - Climate-associated variation in the within-season dynamics of juvenile ticks in California
## Author(s) - S Sambado, A J MacDonald, A Swei, C J Briggs

### Script 3. Data figures


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## upload necessary packages
library(readr)
library(ggplot2)
library(tidyverse)
library(rstatix)
library(MASS)
library(lme4)
library(psych)
library(aod)
library(car)
library(ggpubr)
library(RColorBrewer)
library(ggridges)
library(grid)
library(ggmap)
library(ggrepel)
library(ggspatial)
 
## upload necessary data
# tick data
drag_climate_lifestage <- read.csv("data/drag_climate_lifestage.csv")

liz_climate_lifestage <- read.csv("data/liz_climate_lifestage.csv")

liz_wide_climate_prop <- read.csv("data/liz_wide_climate_prop.csv")

drag_climate_nolifestage <- read.csv("data/drag_climate_nolifestage.csv")

liz_climate_nolifestage <- read.csv("data/liz_climate_nolifestage.csv")

# site info
coords <- read.csv("data/ucnrs_reserve_coords.csv")


## labeler for data
location_names <- c('coastal' = "Coastal",
                    'inland' = "Inland")

region_names <- c("northern" = "Northern", 
                  "central" = "Central",
                  "southern" = "Southern")

```

Format data
```{r}
## correct factor levels

# drag ticks
drag_climate_lifestage$life_stage <- factor(drag_climate_lifestage$life_stage,
                                            levels = c("L", "N", "A"),
                                            labels = c("Larva", "Nymph", "Adult")) 
drag_climate_lifestage$month_name <- factor(drag_climate_lifestage$month_name, 
                                            levels = c("January", "March", "April", "May", "June"))

# lizard ticks
liz_climate_lifestage$month_name <- factor(liz_climate_lifestage$month_name, 
                                            levels = c("March", "April", "May", "June"))

liz_climate_lifestage$region <- factor(liz_climate_lifestage$region,
                                       levels = c("northern", "central", "southern"))  

## make sync data
sync_data <- liz_climate_lifestage %>% 
  filter(count >=0) %>% 
  pivot_wider(names_from = lifestage, values_from = count) %>% 
  group_by(reserve, lizard_num, month_numeric) %>% 
  summarise(larvae = sum(larval, na.rm = T),
            nymphs = sum(nymphal, na.rm = T)) %>% 
  mutate(sync = 
           case_when(
             larvae >0 & nymphs > 0 ~ 1,
             TRUE ~ 0
           ))
sync_data$sync <- factor(sync_data$sync) 

```

# Main Text

## Figure 1

Conceptual Figure. Phenological patterns of juvenile ticks conceptual diagram. Questing activity is represented by the density curves of ticks (blue = nymphs, gold = larvae). The sequential order of life stage emergence within a given season is hypothesized to drive enzootic transmission dynamics of tick-borne pathogens. The panels represent three scenarios which may amplify the transmission of different pathogens such as (a) horizontal transmission of B. burgdorferi, (b) synchronous feeding of A. phagocytophilum, and (c) vertical transmission of B. miyamotoi. 

## Figure 2

Map of study region illustrating the latitudinal and climatic gradients captured by sampling. (a) Study locations are situated throughout the University of California Natural Reserve System with designation of humidity gradient (coastal = teal, inland = brown). We highlight the shift of two important climate variables (b) mean temperature and (c) max vapor pressure deficit (vpd) across the peak tick season (April – June). 

Panel A - make insert of california 
```{r}
# add site abbreviations
ca_coords <- coords %>% 
  dplyr::select(site, latitude, longitude, location) %>% 
  unique() %>% 
  rename(lon = longitude,
         lat = latitude) %>% 
  filter(!is.na(lat)) %>% 
  mutate(abbrev = case_when(site == "McLaughlin" ~ "ML",
                          site == "Quail Ridge" ~ "QR", 
                          site == "Point Reyes" ~ "PR", 
                          site == "Fort Ord" ~ "FO", 
                          site == "Hastings" ~ "HT", 
                          site == "Big Creek" ~ "BC", 
                          site == "Rancho Marino" ~ "RM", 
                          site == "Sedgwick" ~ "SD", 
                          site == "Coal Oil Point" ~ "CO", 
                          site == "Stunt Ranch" ~ "SR", 
                          site == "Santa Cruz Island" ~ "SC", 
                          TRUE ~ site
                          ))

# make targeted coordinates
ca_location <- with(ca_coords,
                    c(lon = median(lon, na.rm = TRUE)+.2,
                      lat = median(lat, na.rm = TRUE)+.3)) 


# register google kep for ggmap
#register_google(key = "AIzaSyC7-ux51sDeZvS8YxNxOkCZ-5_hMYpfAvY", write = TRUE)


# create base map of california
ca_base <- get_map(location = ca_location,
                   zoom = 7, 
                   maptype = "satellite")


# plot CA map with sites
ca_map <- 
ggmap(ca_base) +
  theme(legend.direction = "vertical",
        legend.position = "top",
        legend.key = element_rect(fill = "white"),legend.title.align = .5,
        legend.title = element_text(face = "bold"),
        axis.title = element_text(face = "bold", size = 12),
        title = element_text(face = "bold", size = 14)) +
  guides(color = FALSE, fill = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  #geom_text_repel(data = ca_coords, aes(label = site),
  #                box.padding = .8, max.overlaps = Inf) +
  geom_label_repel(data = ca_coords, aes(label = site),
                   box.padding = .2, label.padding = .12,
                   label.r = .1,
                   max.overlaps = Inf,  size = 4) +
  geom_point(data = ca_coords,
             aes(x = lon, y = lat, fill = location),color = "black", size = 2.5, pch = 21) +
  scale_fill_manual(values = c("#94d2bd", "#ddb892")) +
  annotation_scale(location = "bl", pad_y = unit(0.5,"cm"), text_face = "bold", 
                   bar_cols = c("white", "grey"), text_col = "white") +
  coord_sf(crs = 4326)+
  annotation_north_arrow(location = "bl", pad_y = unit(0.82,"cm")) 

#ggsave(ca_map, file = "figures/ca_map_sites.jpeg", dpi = 600)
```

Panel B & C -- CA climate data

First, download prism data
```{r}
prism_set_dl_dir("~/prismtmp")

# pull from prism()
# get_prism_monthlys(type = "tmean",
#                    year = 2021,
#                    mon = 4:6,
#                    keepZip = FALSE)
# 
# get_prism_monthlys(type = "vpdmax",
#                    year = 2021,
#                    mon = 4:6,
#                    keepZip = FALSE)

# save to space
prism_archive_ls()
pd_to_file(prism_archive_ls())

### FORMAT TMAX
## april
tmean_a <- prism_archive_subset("tmean", "monthly", years = 2021, mon = 4)
ls_prism_data(name = TRUE)
#new_file_a <- c(1)
RS_a_tmean <- prism_stack(ls_prism_data()[1,1]) # change first number 1-6
to_slice_a_tmean <- grep("_202104", RS_a_tmean[,1], value = T)
df_a_tmean <- data.frame(rasterToPoints(RS_a_tmean))
names(df_a_tmean)[1:3] <- c("lon", "lat", "value")
df_a_tmean$month <- "April"
df_a_tmean$type <- "Mean Temperature (C)"

## may
tmean_m <- prism_archive_subset("tmean", "monthly", years = 2021, mon = 5)
ls_prism_data(name = TRUE)
#new_file_a <- c(1)
RS_m_tmean <- prism_stack(ls_prism_data()[2,1]) # change first number 1-6
df_m_tmean <- data.frame(rasterToPoints(RS_m_tmean))
names(df_m_tmean)[1:3] <- c("lon", "lat", "value")
df_m_tmean$month <- "May"
df_m_tmean$type <- "Mean Temperature (C)"

## june
tmean_j <- prism_archive_subset("tmean", "monthly", years = 2021, mon = 6)
ls_prism_data(name = TRUE)
#new_file_a <- c(1)
RS_j_tmean <- prism_stack(ls_prism_data()[3,1]) # change first number 1-6
df_j_tmean <- data.frame(rasterToPoints(RS_j_tmean))
names(df_j_tmean)[1:3] <- c("lon", "lat", "value")
df_j_tmean$month <- "June"
df_j_tmean$type <- "Mean Temperature (C)"

## merge all temp together
tmean_2021 <- rbind(df_a_tmean, df_m_tmean,df_j_tmean)



### FORMAT VPDMAX

## april
vpdmax_a <- prism_archive_subset("vpdmax", "monthly", years = 2021, mon = 4)
ls_prism_data(name = TRUE)
RS_a_vpdmax <- prism_stack(ls_prism_data()[4,1]) # change first number 1-6
df_a_vpdmax <- data.frame(rasterToPoints(RS_a_vpdmax))
names(df_a_vpdmax)[1:3] <- c("lon", "lat", "value")
df_a_vpdmax$month <- "April"
df_a_vpdmax$type <- "Max VPD (kPA)"

ggplot() +
   geom_raster(data = df_a_vpdmax, aes(x = lon, y = lat, fill = value)) +
   theme_nothing()


## may
vpdmax_m <- prism_archive_subset("vpdmax", "monthly", years = 2021, mon = 5)
ls_prism_data(name = TRUE)
#new_file_a <- c(1)
RS_m_vpdmax <- prism_stack(ls_prism_data()[5,1]) # change first number 1-6
df_m_vpdmax <- data.frame(rasterToPoints(RS_m_vpdmax))
names(df_m_vpdmax)[1:3] <- c("lon", "lat", "value")
df_m_vpdmax$month <- "May"
df_m_vpdmax$type <- "Max VPD (kPA)"

## june
vpdmax_j <- prism_archive_subset("vpdmax", "monthly", years = 2021, mon = 6)
ls_prism_data(name = TRUE)
#new_file_a <- c(1)
RS_j_vpdmax <- prism_stack(ls_prism_data()[6,1]) # change first number 1-6
df_j_vpdmax <- data.frame(rasterToPoints(RS_j_vpdmax))
names(df_j_vpdmax)[1:3] <- c("lon", "lat", "value")
df_j_vpdmax$month <- "June"
df_j_vpdmax$type <- "Max VPD (kPA)"

## merge all temp together
vpdmax_2021 <- rbind(df_a_vpdmax, df_m_vpdmax,df_j_vpdmax)

```

Next, make map of climate data
```{r}
### TMEAN 

# set limits
minLat = 33.5 
maxLat = 39.5 
minLon = -123.5 
maxLon = -118 

# filter dataframe to limits
month.df.tmean <- tmean_2021 %>% filter(minLat <lat, lat <maxLat,
                                   minLon<lon, lon < maxLon) %>% 
  mutate(month = factor(month,
                           labels = c("April", "May","June"),
                           levels = c("April", "May","June")))

# make TMEAN map
tmean_2021_map <- 
ggplot() +
  geom_raster(data = month.df.tmean, aes(x = lon, y = lat, fill = value)) +
  theme_nothing(legend = TRUE) +
  scale_fill_distiller(palette = "Spectral")+
  coord_fixed(ratio = 1.3) +
  geom_point(data = ca_coords, aes(x = lon, y = lat), size = 2.5) +
  geom_text(data = ca_coords,aes(x = lon, y = lat, label = abbrev), 
             check_overlap = T, nudge_x = 0.5, nudge_y = .2,
            parse = TRUE, size = 4.5, face = "bold")+
  facet_wrap(~month, ) +
  labs(fill = "Mean Temperature (°C)")+
  theme(legend.position = "bottom",
        legend.title = element_text(face = "bold", vjust = .8),
        strip.background = element_rect(color = "white", fill = "white"),
        strip.text = element_text(face = "bold", size = 14, vjust = -.5))
#ggsave(tmean_2021_map, file = "figures/tmean_2021_map.jpeg", dpi = 600)


### VPD MAX
# set limits
minLat = 33.5 #34.00138
maxLat = 39.5 #38.87289
minLon = -123.5 #-122.7318
maxLon = -118 #-118.656

# filter dataframe to limits
month.df.vpdmax <- vpdmax_2021 %>% filter(minLat <lat, lat <maxLat,
                                   minLon<lon, lon < maxLon) %>% 
  mutate(month = factor(month,
                           labels = c("April", "May","June"),
                           levels = c("April", "May","June")))

# make VPDMAX
vpdmax_2021_map <- 
ggplot() +
  geom_raster(data = month.df.vpdmax, aes(x = lon, y = lat, fill = value)) +
  theme_nothing(legend = TRUE) +
  scale_fill_distiller(palette = "Spectral", limits = c(0,45), na.value = "#d53e4f")+
  coord_fixed(ratio = 1.3) +
  geom_point(data = ca_coords, aes(x = lon, y = lat), size = 2.5) +
  geom_text(data = ca_coords,aes(x = lon, y = lat, label = abbrev), 
             check_overlap = T, nudge_x = 0.5, nudge_y = .2,
            parse = TRUE, size = 4.5, face = "bold")+
  facet_wrap(~month, ) +
  labs(fill = "Max VPD (kPA)")+
  theme(legend.position = "bottom",
        legend.title = element_text(face = "bold", vjust = .8),
        strip.background = element_rect(color = "white", fill = "white"),
        strip.text = element_text(face = "bold", size = 14, vjust = -.5))
#ggsave(vpdmax_2021_map, file = "figures/vpdmax_2021_map.jpeg", dpi = 600)

```


## Figure 3

Abundance patterns for (a) attached and (b) questing I. pacificus ticks vary by collection methods. Attached ticks represent the mean tick burden on an individual lizard per reserve per month. The box spans the first and third quartiles, the median is marked by a thick horizontal line, whisker represent the minimum and maximum values excluding outliers, and solid circles represent outliers. 

```{r}
fig_3_drag <- drag_climate_lifestage %>%
  group_by(reserve, month_name, life_stage) %>%
  summarise(mean = mean(count)) %>%
    filter(month_name != "January") %>%
ggplot(aes(x = month_name, y = mean, fill = life_stage)) +
  geom_boxplot() +
  scale_fill_manual(values = c( "#0077bb", "#ee7733", "#bbbbbb")) +
  theme_bw() +
  labs(y = "Log of Mean Tick Counts", x = "Month", fill = "Life stage", title = "Drag Cloth") + # , tag = "Drag Cloth"
  theme(axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        title = element_text(face = "bold"))  +
  scale_y_log10()


fig_3_liz <- liz_climate_lifestage %>%
  group_by(reserve, month_name, lifestage) %>%
  summarise(mean = mean(count)) %>%
    filter(month_name != "NA") %>%
ggplot(aes(x = month_name, y = mean, fill = lifestage)) +
  geom_boxplot() +
  scale_fill_manual(values = c( "#0077bb", "#ee7733", "#bbbbbb")) +
  theme_bw() +
  labs(y = "Log of Mean Individual Lizard Burden", x = "Month", fill = "Lifestage", title = "Lizard Burden") + #, tag = "Lizard Burden"
  theme(axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12),
        title = element_text(face = "bold")) +
  guides(fill = FALSE) +
  scale_y_log10()



fig_3 <- ggarrange(fig_3_liz +  rremove("xlab") + rremove("ylab"), 
                   fig_3_drag + rremove("ylab"), 
          ncol = 1, 
          common.legend = TRUE, legend = "bottom", 
          align = "hv", labels = c("A", "B"))

fig_3_annotated <- annotate_figure(fig_3, left = text_grob("log10(Mean Tick Counts)", rot = 90, face = "bold", size = 14, just = "center")) 

#ggsave(fig_3_annotated, file = "figures/fig_3_annotated_v6.jpeg")

```



## Figure 4

Regional within-season patterns of juvenile I. pacificus ticks vary by (a) latitudinal (i.e., temperature) and (b) climatic (i.e., humidity) gradients. Upper error-bars (mean + sd) are presented. 

```{r}
# upper error bars
liz_loc <- 
  liz_climate_lifestage %>%
  group_by(month_name, location) %>%
  summarise(mean = mean(count),
            sd = sd(count),
            ymin = mean-sd,
            ymax = mean+sd) %>%
  filter(month_name != "NA") %>%
  ggplot(aes(x = month_name, y = mean, fill = location)) +
  geom_errorbar(aes(ymin = mean, ymax = ymax), color = "grey50",width = .2, size = .8)+
  geom_bar( position = "stack", stat = "identity", color = "grey10") +
  facet_wrap(~location, nrow = 2, scale = "free_y", labeller = as_labeller(location_names)) +
  scale_fill_manual(values = c("#6699cc","#997700")) +
  theme_bw() +
  labs(y = "Mean Burden Counts", x = "Month") + # , title = "(A) Lizard Burden"
    theme(axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12),
        title = element_text(face = "bold"),
        strip.text.x = element_text(face = 'bold', size = 10),
        strip.background = element_rect(fill = 'gray95')) +
  guides(fill = FALSE) 



liz_reg <- 
liz_climate_lifestage %>%
  group_by(month_name, region) %>%
  summarise(mean = mean(count),
            sd = sd(count),
            ymin = mean-sd,
            ymax = mean+sd) %>%
  filter(month_name != "NA") %>%
  ggplot(aes(x = month_name, y = mean, fill = region)) +
  geom_errorbar(aes(ymin = mean, ymax = ymax),color = "grey50", width = .2, size = .8)+
  geom_bar(position = "stack", stat = "identity", color = "grey10") +
  facet_wrap(~region, nrow = 3, scale = "free_y", labeller = as_labeller(region_names)) +
  scale_fill_manual(values = c("#004488","#ddaa33","#bb5566" )) +
  theme_bw() +
  labs(y = "Mean Lizard Burden Counts", x = "Month") +
    theme(axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12),
        title = element_text(face = "bold"),
        strip.text.x = element_text(face = 'bold', size = 10),
        strip.background = element_rect(fill = 'gray95')) +
  guides(fill = FALSE)

fig_4_annotated <- ggarrange(liz_reg + rremove("xlab"), liz_loc + rremove("ylab") + rremove("xlab"),
          align = "hv",
           labels = c("A", "B"))

#ggsave(fig_4_annotated, file = "figures/fig_4_annotated_v3.jpeg")
```



## Figure 5
Individual reserves, which represent a unique combination of temperature and humidity, vary in their within-season dynamics of juvenile I. pacificus from attached lizards. Within-season dynamics such as (a) duration of activity, (b) emergence rate, (c) proportion of synchrony events, and (d) sequential pattern are different for each reserve. (a) Density curves represent the monthly abundance of juvenile ticks (top graph = larvae, bottom graph = nymphs) during the sampling season (March-June) at each reserve. (b) Cumulative proportion represents the emergence rate of juvenile ticks by reserve during the sampling season. Color of the reserves represent the descending order of mean temperature in March for each reserve (i.e., highest mean temperature in March was CO, lowest was FO). (c) The proportion of synchronous events (i.e. when both larva and nymph feed on an individual lizard) at each reserve during the sampling season. (d) The sequential order of life stages emerging (i = larvae first, ii = synchrony, iii = nymphs first).

Panel A - Duration

```{r}
## organize data
 liz_climate_lifestage2 <- liz_climate_lifestage

liz_climate_lifestage2$reserve <- factor(liz_climate_lifestage2$reserve,
                        levels = c("McLaughlin", "Quail Ridge", "Point Reyes",
                                   "Fort Ord", "Hastings", 
                                   "Big Creek", "Rancho Marino",
                                    "Santa Cruz Island",
                                  "Sedgwick", "Coal Oil Point",
                                   "Stunt Ranch"))
 

# plot
geom_ridge_larvae <-  
liz_climate_lifestage2 %>%
    filter(lifestage == "larval") %>% 
  drop_na() %>% 
  uncount(count) %>% 
  filter(reserve != "NA" ) %>% #& lifestage =="Nymphs"
  add_row(reserve = "Coal Oil Point") %>% 
  mutate(reserve = factor(reserve,
                      levels = c("McLaughlin", "Quail Ridge", "Point Reyes",
                                 "Fort Ord", "Hastings", 
                                 "Big Creek", "Rancho Marino",
                                  "Santa Cruz Island",
                                 "Sedgwick", "Coal Oil Point",
                                 "Stunt Ranch"))) %>% 
  ggplot(aes(x = month_numeric, y = fct_rev(reserve), fill = reserve, group = reserve)) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      bandwidth = .35,
                      rel_min_height = .0001,
                      alpha = .99,
                      calc_ecdf = TRUE,
                      #scale = 1.3,
                      jittered_points = TRUE,
                      position = position_points_jitter(width = .1, height =0),
                       point_shape = "|", 
                       point_size = 1, point_alpha = .7, alpha = .5) +
  theme_bw() +
  labs(x = "", y = "") +
  scale_x_discrete(limits = c(3,4,5,6),
                     labels = c("M","A", "M", "J"),
                   expand = c(0.01,0)) +
  scale_fill_viridis_d()+
    #scale_fill_viridis(discrete = TRUE, option = "D") +
  theme(        strip.text.x = element_text(face = 'bold', size = 10),
        strip.background = element_rect(fill = 'gray95'),
        strip.text.y = element_text(face = 'bold', size = 10),
        axis.text.x = element_text(face = 'bold', size = 16),
        axis.text.y = element_text(face = 'bold', size = 16),
        axis.title = element_text(face = "bold", size = 16),
        axis.text = element_text(face = "bold", size = 14)) +
  guides(fill = FALSE) +
      scale_y_discrete(expand = expand_scale(mult = c(0.01, .13))) +
  scale_y_discrete(labels = c("McLaughlin" = "ML",
                              "Quail Ridge" = "QR",
                              "Point Reyes" = "PR",
                              "Fort Ord" = "FO",
                              "Hastings" = "HT",
                              "Big Creek" = "BC",
                              "Rancho Marino" = "RM",
                              "Coal Oil Point" = "CO",
                              "Sedgwick" = "SD",
                              "Santa Cruz Island" = "SC",
                              "Stunt Ranch" = "SR"))




geom_ridge_larvae_an <- geom_ridge_larvae + annotate("text", x = 2.8, y = 10.5, label = "Larvae", fontface = 2, size = 7)



######### NYMPH

geom_ridge_nymph <-  
liz_climate_lifestage2 %>%
    filter(lifestage == "nymphal") %>% 
  drop_na() %>% 
  uncount(count) %>% 
  filter(reserve != "NA" ) %>% #& lifestage =="Nymphs"
  ggplot(aes(x = month_numeric, y = fct_rev(reserve), fill = reserve, group = reserve)) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      bandwidth = .35,
                      rel_min_height = .0001,
                      alpha = .99,
                      calc_ecdf = TRUE,
                      #scale = 1.3,
                      jittered_points = TRUE,
                      position = position_points_jitter(width = .1, height =0),
                       point_shape = "|", 
                       point_size = 1, point_alpha = .7, alpha = .5) + 
    scale_fill_viridis_d() +
  theme_bw() +
  labs(x = "", y = "") +
  scale_x_discrete(limits = c(3,4,5,6),
                     labels = c("M","A", "M", "J"),
                   expand = c(0.01,0)) +
  theme(
        strip.text.x = element_text(face = 'bold', size = 10),
        strip.background = element_rect(fill = 'gray95'),
        strip.text.y = element_text(face = 'bold', size = 10),
        axis.text.x = element_text(face = 'bold', size = 16),
        axis.text.y = element_text(face = 'bold', size = 16),
        axis.title = element_text(face = "bold", size = 16),
        axis.text = element_text(face = "bold", size = 14)
        ) +
  guides(fill = FALSE) +
      scale_y_discrete(expand = expand_scale(mult = c(0.01, .13))) +
  scale_y_discrete(labels = c("McLaughlin" = "ML",
                              "Quail Ridge" = "QR",
                              "Point Reyes" = "PR",
                              "Fort Ord" = "FO",
                              "Hastings" = "HT",
                              "Big Creek" = "BC",
                              "Rancho Marino" = "RM",
                              "Coal Oil Point" = "CO",
                              "Sedgwick" = "SD",
                              "Santa Cruz Island" = "SC",
                              "Stunt Ranch" = "SR"))

geom_ridge_nymph_an <- geom_ridge_nymph + annotate("text", x = 2.8, y = 10.5, label = "Nymphs", fontface = 2, size = 7)

  
geom_ridge_nymph_an2 <- ggarrange(geom_ridge_nymph_an, labels = "", width = .8)
geom_ridge_larvae_an2 <- ggarrange(geom_ridge_larvae_an, labels = "", width = .8)
#ggsave(geom_ridge_nymph_an, file = "figures/fig5_a3.jpeg" )
#ggsave(geom_ridge_larvae_an, file = "figures/fig5_a3_larvae.jpeg")
```

Panel B - emergence
```{r}
## emergence
liz_climate_lifestage3 <- liz_climate_lifestage
liz_climate_lifestage3$reserve <- factor(liz_climate_lifestage3$reserve, 
                                         levels = c("Coal Oil Point", 
                                                    "Sedgwick", "Stunt Ranch",
                                                    "McLaughlin", "Quail Ridge", 
                                                    "Rancho Marino", "Hastings",
                                                    "Big Creek", "Fort Ord", 
                                                    "Santa Cruz Island"))

emergence <- 
  liz_climate_lifestage3 %>% 
  filter(reserve != "NA" & reserve != "Santa Cruz Island") %>% 
  group_by(reserve, month_numeric) %>%
  summarise(month_tot = sum(count)) %>% 
  mutate(cum = cumsum(month_tot)/sum(month_tot)) %>% 
  ggplot(aes(x = month_numeric, y = cum, color = reserve)) +
    geom_hline(yintercept = .50) +
  geom_line(size = 1.5)+ 
  scale_color_manual(values = c("Coal Oil Point" = "#780116",
                                "Sedgwick" = "#C32f27",
                                "Stunt Ranch" = "#F94144",
                                "McLaughlin" = "#F3722C",
                                "Quail Ridge" = "#F9C74F",
                                "Rancho Marino" = "#90BE6D",
                                "Hastings" = "#43AA8B",
                                "Big Creek" = "#4D908E",
                                "Fort Ord" = "#577590"
                                ),
                     labels = c("McLaughlin" = "ML",
                              "Quail Ridge" = "QR",
                              "Point Reyes" = "PR",
                              "Fort Ord" = "FO (coolest)",
                              "Hastings" = "HT",
                              "Big Creek" = "BC",
                              "Rancho Marino" = "RM",
                              "Coal Oil Point" = "CO (hotest)",
                              "Sedgwick" = "SD",
                              "Santa Cruz Island" = "SC",
                              "Stunt Ranch" = "SR")) +
  theme_bw() +
  labs(x = "", y = "Cumulative Proportion", color = "Reserve") +
  theme(axis.title = element_text(face = "bold", size = 16),
        strip.text.y = element_text(face = 'bold', size = 10),
        #axis.text.x = element_text( size = 16, face = "bold"), #face = 'bold',
        legend.title = element_text(face = "bold",size = 13),
        #legend.position = "bottom",
        legend.position = c(.85,.35),
        legend.text = element_text(size = 14),
        legend.margin = unit(-.5,"cm"),
          #axis.title = element_text(face = "bold", size = 16),
        axis.text = element_text(face = "bold", size = 14)) +
  scale_x_continuous(labels = c("3" = "M", "4" = "A", "5" = "M", "6" = "J")) 


## save image
emergence_an <- ggarrange(emergence, labels = "")
#ggsave(emergence_an, file = "figures/fig5_b_v3.jpeg")
```

Panel C - synchrony
```{r}
synchrony <- sync_data %>% 
  mutate(reserve = factor(reserve,
                      levels = c("McLaughlin", "Quail Ridge", "Point Reyes",
                                 "Fort Ord", "Hastings", 
                                 "Big Creek", "Rancho Marino",
                                  "Santa Cruz Island",
                                 "Sedgwick", "Coal Oil Point",
                                 "Stunt Ranch"))) %>% 
ggplot( aes(y = fct_rev(reserve), fill = sync))+
  geom_bar(position = "fill")  +
  theme_classic() + 
  labs(y = "", x = "Proportion of Synchronous Events") +
  scale_fill_manual(name = "Synchronous\n Event", 
                        labels = c("No", "Yes"),
                    values = c("grey90", "#dc050c"))+
  theme(
        strip.text.x = element_text(face = 'bold', size = 16),
        strip.background = element_rect(fill = 'gray95'),
        strip.text.y = element_text(face = 'bold', size = 10),
        axis.text.x = element_text(face = 'bold', size = 12),
        axis.text.y = element_text(face = 'bold', size = 16),
        legend.title = element_text(face = 'bold'),
        axis.title = element_text(face = "bold", size = 15),
        legend.direction = "horizontal",
        legend.text = element_text(size = 13),
        legend.position = c(.75,.1)
        ) +
  geom_vline(xintercept = .50) +
  scale_x_continuous(expand = c(0.02,0))  +
  #scale_fill_manual(values = c("grey90", "#dc050c")) +
    scale_y_discrete(labels = c("McLaughlin" = "ML",
                              "Quail Ridge" = "QR",
                              "Point Reyes" = "PR",
                              "Fort Ord" = "FO",
                              "Hastings" = "HT",
                              "Big Creek" = "BC",
                              "Rancho Marino" = "RM",
                              "Coal Oil Point" = "CO",
                              "Sedgwick" = "SD",
                              "Santa Cruz Island" = "SC",
                              "Stunt Ranch" = "SR")) 

synchrony_an <- ggarrange(synchrony, labels = "")
#ggsave(synchrony_an, file = "figures/fig5_c.jpeg")
```

Panel D - phenology patterns
```{r}
drag_sub <- drag_climate_lifestage %>% 
  dplyr::select(reserve, count, life_stage, month_name) %>% 
  mutate(method = "drag") %>% 
  mutate(across(c(life_stage), factor)) 

levels(drag_sub$life_stage) <- list("adult" = "A", "nymphal" = "N", "larval" = "L")


liz_sub <- liz_climate_lifestage %>% 
  dplyr::select(reserve, count, lifestage, month_name) %>% 
  rename(life_stage = lifestage) %>% 
  mutate(method = "lizard")

full_sub <- rbind(drag_sub, liz_sub)


full_sub$month_name <- factor(full_sub$month_name, 
                                            levels = c("January", "March", "April", "May", "June"))



full_sub$reserve <- factor(full_sub$reserve,
                      levels = c("McLaughlin", "Quail Ridge", "Point Reyes",
                                 "Fort Ord", "Hastings",
                                 "Big Creek", "Rancho Marino",
                                  "Santa Cruz Island",
                                 "Sedgwick", "Coal Oil Point",
                                 "Stunt Ranch"))


labels = c("McLaughlin" = "ML (ii)",
                              "Quail Ridge" = "QR (i)",
                              "Point Reyes" = "PR (i)",
                              "Fort Ord" = "FO (ii)",
                              "Hastings" = "HT (i)",
                              "Big Creek" = "BC (ii)",
                              "Rancho Marino" = "RM (ii)",
                              "Coal Oil Point" = "CO (i)",
                              "Sedgwick" = "SD (ii)",
                              "Santa Cruz Island" = "SC (ii)",
                              "Stunt Ranch" = "SR (ii)")

lab_method <- c("drag" = "Drag", "lizard" = "Lizard")
full_sub %>% 
  filter(life_stage != "adult") %>% 
  ggplot(aes(x = month_name, y = count, fill = life_stage)) +
  geom_bar(stat = "identity") +
  facet_wrap(~reserve)


full_sub2 <- full_sub



full_sub2$life_stage <- factor(full_sub2$life_stage,
                               levels = c("larval", "nymphal", "adult"),
                               labels = c("Larvae", "Nymphs", "Adults"))

facet_grid_mthod_reserve <- 
  full_sub2 %>% 
  filter(month_name == "January" |
         month_name == "March" |
           month_name == "April" |
           month_name == "May" |
           month_name == "June") %>%  
  group_by(reserve, month_name, life_stage) %>% #method, 
  summarise(total = sum(count)) %>% 
    #mutate(total = total+1) %>% # for log_10 
    filter(month_name != "January") %>% 
  ggplot(aes(x = month_name, y = total,fill = life_stage , group = reserve)) + #, fill = life_stage , group = reserve
  geom_col(stat = "identity", position = "fill") + #stat = "identity", 
  scale_fill_manual(values = c( "#0077bb", "#ee7733","#bbbbbb")) +
  #facet_grid(method~reserve, scales = "free_y", labeller = labeller(reserve = labels, method = lab_method)) +
  facet_wrap(~reserve) +
  theme_bw() +
  theme(legend.position = "right",#bottom
        strip.background = element_rect(fill = "grey95"),
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(#angle = 45, 
          #hjust = 1, 
          size = 8),
        axis.title = element_text(size = 12, face = "bold")) +
  labs(y = "Proportion", x = "", fill = "Life stage")  +
  scale_x_discrete(labels = c("January" = "J",
                              "March" = "M",
                              "April" = "A",
                              "May" = "M",
                              "June" = "J")) +
  guides(fill = FALSE)# +
  scale_y_log10() 

facet_grid_mthod_reserve_an <- ggarrange(facet_grid_mthod_reserve, labels = "")
#ggsave(facet_grid_mthod_reserve_an, file = "Revisions/fig5_d_v4.jpeg", height = 4)

```

